---
title: "class3DTree"
author: "Suhail AK"
date: "May 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Decision Tree

In LR -> sales = m.TV + C

so in DT

sales = 5.5 if TV < 200 else 100

in DT if condition passes it will be in RHS
else LHS
widely used method CART(classification and regression trees)

```{r}
library(dplyr)
library(ggplot2)
library(tree)
library(recommenderlab)


adv <- read.csv("Advertising.csv")
View(adv)

model <- tree(sales~TV,data=adv)
{{plot(model)
  text(model)}}


```
- Always root node will have all the sample in this case(200 rows)

```{r}
nrow(adv) #which is 200

adv %>% filter(TV<30.05) %>% summarise(val=n(),avg=mean(sales))#here for 26 samples are there in TV, which has value less than 30.05
range(adv$TV)
aa <- adv %>% arrange(TV)

View(aa)

```


- to calculate error for Decision TREE for regression problem
```{r}
cut <- 100
df <- data.frame(x=adv$TV,y=adv$sales)

sample_left <- df %>% filter(x<cut)
sample_right <- df %>% filter(x>cut)

nrow(sample_left)

avg_left <- mean(sample_left$y)
avg_right <- mean(sample_right$y)
View(df)

df$yhat <- ifelse(df$x<cut,avg_left,avg_right)
View(df)
mse <- sum((df$y-df$yhat)^2)/nrow(df)
mse

```

class4 starts from here Decision TREE

analyzing the decision tree by finding each cut as shown in the tree function.


```{r}
df <- data.frame(x=adv$TV,y=adv$sales)
tv_uniq <- sort(unique(df$x)) #its better take a cut when its sorted
View(df)
cuts <- (tv_uniq[1:length(tv_uniq)-1] + tv_uniq[2:length(tv_uniq)]) /2
length(cuts)
mse_cuts <- c()

for (cut in cuts) {
  
  sample_left <- df %>% filter(x<cut)
  sample_right <- df %>% filter(x>cut)
  avg_left <- mean(sample_left$y)
  avg_right <- mean(sample_right$y)
  
  
  df$yhat <- ifelse(df$x<cut,avg_left,avg_right)
  curr_mse <- sum((df$y-df$yhat)^2)/nrow(df)
  mse_cuts <- c(mse_cuts,curr_mse)
  
}
result <- data.frame(cuts=cuts,mse=mse_cuts)
View(result)

result %>% arrange(mse) %>% head(1) #here this cut is after the root node



```

now for 83 samples ie tv <122.05
i.e 30.05 is the cut

```{r}
temp <- adv %>% filter(TV<122.05)
df <- data.frame(x=temp$TV,y=temp$sales)
tv_uniq <- sort(unique(df$x)) 
cuts <- (tv_uniq[1:length(tv_uniq)-1] + tv_uniq[2:length(tv_uniq)]) /2
length(cuts)
mse_cuts <- c()

for (cut in cuts) {
  
  sample_left <- df %>% filter(x<cut)
  sample_right <- df %>% filter(x>cut)
  avg_left <- mean(sample_left$y)
  avg_right <- mean(sample_right$y)
  
  
  df$yhat <- ifelse(df$x<cut,avg_left,avg_right)
  curr_mse <- sum((df$y-df$yhat)^2)/nrow(df)
  mse_cuts <- c(mse_cuts,curr_mse)
  
}
result <- data.frame(cuts=cuts,mse=mse_cuts)
View(result)

result %>% arrange(mse) %>% head(1)

```

cuts for 240.9

```{r}
temp <- adv %>% filter(TV>122.05)
df <- data.frame(x=temp$TV,y=temp$sales)
tv_uniq <- sort(unique(df$x)) 
cuts <- (tv_uniq[1:length(tv_uniq)-1] + tv_uniq[2:length(tv_uniq)]) /2
length(cuts)
mse_cuts <- c()

for (cut in cuts) {
  
  sample_left <- df %>% filter(x<cut)
  sample_right <- df %>% filter(x>cut)
  avg_left <- mean(sample_left$y)
  avg_right <- mean(sample_right$y)
  
  
  df$yhat <- ifelse(df$x<cut,avg_left,avg_right)
  curr_mse <- sum((df$y-df$yhat)^2)/nrow(df)
  mse_cuts <- c(mse_cuts,curr_mse)
  
}
result <- data.frame(cuts=cuts,mse=mse_cuts)
View(result)

result %>% arrange(mse) %>% head(1)
```


for multiple column in decision tree

```{r}
model2 <- tree(sales~.,data = adv)
model2

{{plot(model2)
  text(model2)}}

```

to find the first cut i.e is TV at 122.05

```{r}
temp <- adv
df <- data.frame(TV=temp$TV,radio=temp$radio,y=temp$sales)

tv_uniq <- sort(unique(df$TV)) 
tv_cuts <- (tv_uniq[1:length(tv_uniq)-1] + tv_uniq[2:length(tv_uniq)]) /2
length(tv_cuts)

radio_uniq <- sort(unique(df$radio)) 
radio_cuts <- (radio_uniq[1:length(radio_uniq)-1] + radio_uniq[2:length(radio_uniq)]) /2
length(radio_cuts)

predictors=c(rep("TV",length(tv_cuts)),
             rep("radio",length(radio_cuts)))

cuts <- c(tv_cuts,radio_cuts)
mse_cuts <- c()

for (i in seq(1,length(cuts))) {
  cut=cuts[i]
  samples_left <- df[df[,predictors[i]]<cut,]
  samples_right <- df[df[,predictors[i]]>cut,]
  avg_left <- mean(samples_left$y)
  avg_right <- mean(samples_right$y)
  
  
  
  df$yhat <- ifelse(df[,predictors[i]]<cut,avg_left,avg_right)
  curr_mse <- sum((df$y-df$yhat)^2)/nrow(df)
  mse_cuts <- c(mse_cuts,curr_mse)

}
result <- data.frame(predictor=predictors,cuts=cuts,mse=mse_cuts)
View(result)

result %>% arrange(mse) %>% head(1)


```

to find the second cut i.e for TV > 122.05 which will be radio with 26.85 as a cut

```{r}


temp <- adv %>% filter(TV>122.05)
df <- data.frame(TV=temp$TV,radio=temp$radio,y=temp$sales)

tv_uniq <- sort(unique(df$TV)) 
tv_cuts <- (tv_uniq[1:length(tv_uniq)-1] + tv_uniq[2:length(tv_uniq)]) /2
length(tv_cuts)

radio_uniq <- sort(unique(df$radio)) 
radio_cuts <- (radio_uniq[1:length(radio_uniq)-1] + radio_uniq[2:length(radio_uniq)]) /2
length(radio_cuts)

predictors=c(rep("TV",length(tv_cuts)),
             rep("radio",length(radio_cuts)))

cuts <- c(tv_cuts,radio_cuts)
mse_cuts <- c()

for (i in seq(1,length(cuts))) {
  cut=cuts[i]
  samples_left <- df[df[,predictors[i]]<cut,]
  samples_right <- df[df[,predictors[i]]>cut,]
  avg_left <- mean(samples_left$y)
  avg_right <- mean(samples_right$y)
  
  
  
  df$yhat <- ifelse(df[,predictors[i]]<cut,avg_left,avg_right)
  curr_mse <- sum((df$y-df$yhat)^2)/nrow(df)
  mse_cuts <- c(mse_cuts,curr_mse)

}
result <- data.frame(predictor=predictors,cuts=cuts,mse=mse_cuts)
View(result)

result %>% arrange(mse) %>% head(1)




```

to find the next cut ie. TV as 194.55 as a cut.

```{r}

temp <- adv %>% filter(TV>122.05,radio > 26.85)
df <- data.frame(TV=temp$TV,radio=temp$radio,y=temp$sales)

tv_uniq <- sort(unique(df$TV)) 
tv_cuts <- (tv_uniq[1:length(tv_uniq)-1] + tv_uniq[2:length(tv_uniq)]) /2
length(tv_cuts)

radio_uniq <- sort(unique(df$radio)) 
radio_cuts <- (radio_uniq[1:length(radio_uniq)-1] + radio_uniq[2:length(radio_uniq)]) /2
length(radio_cuts)

predictors=c(rep("TV",length(tv_cuts)),
             rep("radio",length(radio_cuts)))

cuts <- c(tv_cuts,radio_cuts)
mse_cuts <- c()

for (i in seq(1,length(cuts))) {
  cut=cuts[i]
  samples_left <- df[df[,predictors[i]]<cut,]
  samples_right <- df[df[,predictors[i]]>cut,]
  avg_left <- mean(samples_left$y)
  avg_right <- mean(samples_right$y)
  
  
  
  df$yhat <- ifelse(df[,predictors[i]]<cut,avg_left,avg_right)
  curr_mse <- sum((df$y-df$yhat)^2)/nrow(df)
  mse_cuts <- c(mse_cuts,curr_mse)

}
result <- data.frame(predictor=predictors,cuts=cuts,mse=mse_cuts)
View(result)

result %>% arrange(mse) %>% head(1)


```

Instead of MSE lets calculate variance deviance cutt


```{r}
temp <- adv
df <- data.frame(TV=temp$TV,radio=temp$radio,y=temp$sales)

tv_uniq <- sort(unique(df$TV)) 
tv_cuts <- (tv_uniq[1:length(tv_uniq)-1] + tv_uniq[2:length(tv_uniq)]) /2
length(tv_cuts)

radio_uniq <- sort(unique(df$radio)) 
radio_cuts <- (radio_uniq[1:length(radio_uniq)-1] + radio_uniq[2:length(radio_uniq)]) /2
length(radio_cuts)

predictors=c(rep("TV",length(tv_cuts)),
             rep("radio",length(radio_cuts)))

cuts <- c(tv_cuts,radio_cuts)
vardev_cuts <- c() #variance deviance cut

for (i in seq(1,length(cuts))) {
  cut=cuts[i]
  samples_left <- df[df[,predictors[i]]<cut,]
  samples_right <- df[df[,predictors[i]]>cut,]
  avg_left <- mean(samples_left$y)
  avg_right <- mean(samples_right$y)
  
  
  
  df$yhat <- ifelse(df[,predictors[i]]<cut,avg_left,avg_right)
  curr_vardev <- var(df$y) - (nrow(samples_left)/nrow(df)) * var(samples_left$y) -
    (nrow(samples_right)/nrow(df)) * var(samples_right$y)
  vardev_cuts <- c(vardev_cuts,curr_vardev)

}
result <- data.frame(predictor=predictors,cuts=cuts,vardeviation=vardev_cuts)
View(result)

result %>% arrange(-vardev_cuts) %>% head(1)


```







