---
title: "class5ml"
author: "Suhail AK"
date: "May 14, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Gini Impurities

```{r }
library(ISLR)
library(dplyr)
library(tree)

hr = read.csv('HR Analytics.csv')

dim(hr)
hr_train = hr[1:(0.7*nrow(hr)),]
hr_test = hr[(0.7*nrow(hr)+1):nrow(hr),]

model<- tree(Attrition~.,data= hr_train)
{{plot(model)
  text(model)}}

```

```{r }
#install.packages("rattle")
library(rattle)
library(rpart)

```

```{r }
m1<- rpart(Attrition~OverTime+Gender,data= hr_train)
View(hr_train)
hr_train %>% filter(OverTime=="No" & Attrition==1) %>% nrow()
hr_train %>% filter(OverTime=="Yes" & Attrition==1) %>% nrow()


{{plot(m1)
  text(m1)}}
fancyRpartPlot(m1)
```

#########GINI IMPURITY ###################
###########Input variable: categorical with two classes

```{r }
nrow(hr_train)
left_overtime = hr_train %>% filter(OverTime=='Yes')
right_overtime = hr_train %>% filter(OverTime=='No')
nrow(left_overtime)
nrow(right_overtime)
```

```{r }
library(data.table)
table(left_overtime$Attrition)
GIl = 1-(85/292)^2-(207/292)^2
table(right_overtime$Attrition)
GIr = 1-(75/737)^2-(662/737)^2
GI = (303/1029)*GIl + (726/1029)*GIr
GI
```

```{r }
left_gender = hr_train %>% filter(Gender=='Male')
right_gender = hr_train %>% filter(Gender=='Female')
nrow(left_gender)
nrow(right_gender)
```

```{r }
table(left_gender$Attrition)
table(right_gender$Attrition)
GIL = 1-(96/615)^2 - (519/615)^2
GIR = 1-(64/414)^2 - (350/414)^2
GI2 = nrow(left_gender)/nrow(hr_train)*GIL+nrow(right_gender)/nrow(hr_train)*GIR
GI2
```

```{r }
table(hr_train$Attrition)
1-(869/nrow(hr_train))^2 - (160/nrow(hr_train))^2
```

########### Taking marital status as the input col and splitting it in two parts########

```{r }
marital_status_uniq = unique(hr_train$MaritalStatus)

for (status in marital_status_uniq){
  samples_left = hr_train %>% filter(MaritalStatus==status)
  samples_right = hr_train %>% filter(MaritalStatus !=status)
  
  p0_left = nrow(samples_left %>% filter(Attrition==0))/nrow(samples_left)
  p1_left = nrow(samples_left %>% filter(Attrition==1))/nrow(samples_left)
  gi_left = 1- p0_left^2 - p1_left^2
  
  p0_right = nrow(samples_right %>% filter(Attrition==0))/nrow(samples_right)
  p1_right = nrow(samples_right %>% filter(Attrition==1))/nrow(samples_right)
  gi_right = 1- p0_right^2 - p1_right^2
  
  gi_status = nrow(samples_left)/nrow(hr_train)*gi_left +
    nrow(samples_right)/nrow(hr_train)*gi_right
 
  temp = marital_status_uniq[marital_status_uniq!=status]
  print('Left node')
  print(status)
  print('Right node')
  print(temp)
  print(gi_status)
  print("----------------------------------")
}
                             
```
################ Job column with more than three classes ######################

```{r }
unique(hr_train$JobRole)
```

```{r }
x= c('a','b','c','d')
combn(x,2,simplify = F)
```

```{r }
jobs_uniq = unique(hr_train$JobRole)

combination_left = c()
combination_right = c()
gi_all = c()
for (n in c(1,2,3,4)){
  comb_n = combn(jobs_uniq,n,simplify = F)
  for (i in seq(1,length(comb_n))){
    comb_left = comb_n[[i]]
    comb_right = jobs_uniq[!jobs_uniq %in% comb_left]
    
    samples_left = hr_train %>% filter(JobRole %in% comb_left)
    samples_right = hr_train %>% filter(JobRole %in% comb_right)
    
    p0_left = nrow(samples_left %>% filter(Attrition==0))/nrow(samples_left)
    p1_left = nrow(samples_left %>% filter(Attrition==1))/nrow(samples_left)
    gi_left = 1- p0_left^2 - p1_left^2
    
    p0_right = nrow(samples_right %>% filter(Attrition==0))/nrow(samples_right)
    p1_right = nrow(samples_right %>% filter(Attrition==1))/nrow(samples_right)
    gi_right = 1- p0_right^2 - p1_right^2
    
    gi_role = nrow(samples_left)/nrow(hr_train)*gi_left +
      nrow(samples_right)/nrow(hr_train)*gi_right
    
    combination_left = c(combination_left, paste0(comb_left,collapse = ','))
    combination_right = c(combination_right, paste0(comb_right,collapse = ','))
    gi_all = c(gi_all, gi_role)
  }
}
result= data.frame(left= combination_left,right = combination_right, gi=gi_all)
result%>% arrange(gi)%>%head(1)
View(result)

```

```{r }

model = tree(Attrition~JobRole, data= hr_train)
levels(hr_train$JobRole)
{{plot(model)
  text(model)}}

```

```{r, fig.width=8 }
model = rpart(Attrition~JobRole, data= hr_train)
fancyRpartPlot(model)
model
```

################ Numerical column Input #############################

```{r }

mi_uniq = sort(unique(hr_train$MonthlyIncome))

length(mi_uniq)
cuts_mi = (mi_uniq[1:length(mi_uniq)-1] + 
             mi_uniq[2:length(mi_uniq)])/2
length(cuts_mi)
cuts= c()
gi_all = c()
for( cut in cuts_mi){
  samples_left = hr_train %>% filter(MonthlyIncome < cut)
  samples_right = hr_train %>% filter(MonthlyIncome > cut)
  
    p0_left = nrow(samples_left %>% filter(Attrition==0))/nrow(samples_left)
    p1_left = nrow(samples_left %>% filter(Attrition==1))/nrow(samples_left)
    gi_left = 1- p0_left^2 - p1_left^2
    
    p0_right = nrow(samples_right %>% filter(Attrition==0))/nrow(samples_right)
    p1_right = nrow(samples_right %>% filter(Attrition==1))/nrow(samples_right)
    gi_right = 1- p0_right^2 - p1_right^2
    
    gi_income = nrow(samples_left)/nrow(hr_train)*gi_left +
      nrow(samples_right)/nrow(hr_train)*gi_right
    cuts = c(cuts,cut)
    gi_all = c(gi_all,gi_income)
    
}
 result= data.frame(cuts,gi_all)
 View(result)
result%>% arrange(gi_all)%>%head(1)

```

```{r }
model = rpart(Attrition~MonthlyIncome, data= hr_train)
fancyRpartPlot(model)
model
```

```{r }
model = rpart(Attrition~ OverTime, data= hr_train)
fancyRpartPlot(model)
model
```
###Validate the values in above plot by calculations

```{r }
table(hr_train$Attrition)
samples_left = hr_train %>% filter(OverTime=='No')
samples_right = hr_train %>% filter(OverTime=='Yes')
nrow(samples_left)
nrow(samples_left)/ nrow(hr_train)

nrow(samples_left %>% filter(Attrition==1))/nrow(samples_left)
nrow(samples_right %>% filter(Attrition==1))/nrow(samples_right)
```

```{r }
nrow(hr_train %>% filter(Attrition==0))/nrow(hr_train)
nrow(hr_train %>% filter(Attrition==1))/nrow(hr_train)
```

















## More numerical columns as inputs 

```{r}
library(rpart)
library(dplyr)
library(ggplot2)
library(rattle)
hr <- read.csv("HR Analytics.csv")
View(hr)

model <- rpart(Attrition~MonthlyIncome+Age,data = hr)
fancyRpartPlot(model)

#cp is complexity parameter

model <- rpart(Attrition~MonthlyIncome+Age,data = hr,cp=0.03)
fancyRpartPlot(model)

```

```{r}
#using single numeric column as feature selection

model <- rpart(Attrition~MonthlyIncome,data = hr)
fancyRpartPlot(model)


```






#more than 3 factors in a column 
eg: Marital Status has single,married, div
multiclass input

```{r}
model <- rpart(Attrition~MaritalStatus, data = hr)
fancyRpartPlot(model)

```

```{r}
ms_uniqs <- unique(hr$MaritalStatus)
N <- nrow(hr)
GIs <- c()

for(status in ms_uniqs){
  samples_left <- hr %>% filter(MaritalStatus == status)
  samples_right <- hr %>% filter(MaritalStatus != status)
  
  Nl <- nrow(samples_left)
  Nr <- nrow(samples_right)
  
  #left side gi
  GI_left <- 1- (nrow(samples_left %>% filter(Attrition==0))/Nl)^2 -
    (nrow(samples_left %>% filter(Attrition==1))/Nl)^2
  
  #right side gi
  
  GI_right <- 1- (nrow(samples_right %>% filter(Attrition==0))/Nr)^2 -
    (nrow(samples_left %>% filter(Attrition==1))/Nr)^2
  
  GI <- (Nl(N*GI_left)) + (Nr/N*GI_right)
  GIs <- c(GIs,GI)
  
}
  
result = data.frame(left=ms_uniqs,GI=GIs) 

  
```




```{r}

job_uniqs <- unique(hr$JobRole)

val <- combn(job_uniqs,1,simplify = F)

for (i in val){
  jobs_left <- i[[1]]
  print(jobs_left)
}


```




