---
title: "class3ML"
author: "Suhail AK"
date: "May 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


to find gradient descent




```{r}

x <- rnorm(100)

y=0.05 * x + 50
df <- data.frame(x=x,y=y)
View(df)

lm(y~x,data = df)

```
here we got x as 0.05
and x intercept is almost 0.
now using gradient descent we will arrive at 0.05(which is a slope)

gradient descent: reduce the error
gradient ascent:  increase the accuracy




```{r}

m <- 100
alpha <- 0.1 #keep changing alpha values
iterations <- 1000
errors_vals <- c()
for (i in seq(1,iterations)){
  df <- mutate(df,mx_vals = m * x)
  df <- mutate(df, y_mx_vals = (y-mx_vals)^2)
  cur_error <- sum(df$y_mx_vals)/nrow(df)
  errors_vals <- c(errors_vals,cur_error)
  df =  dplyr::mutate(df, xy_vals = x * y)
  df = dplyr::mutate(df, mx_square = m * (x^2))
  df = dplyr::mutate(df, xy_minus_mx2 = xy_vals - mx_square)
  m_gradient = -2/nrow(df) * sum(df$xy_minus_mx2)
  m = m - alpha * m_gradient
  
}
print(m)
View(df)

{{plot(errors_vals)
  lines(m,type = "l",col="red")}}


```

to reduce the iteration we have to give up on precision



gradient descent example


```{r}


x <- rnorm(100)
y=(0.05 * x) + 100
df <- data.frame(x=x,y=y)

lm(y~x,data = df)


```

```{r}
m <- 0
alpha <- 0.01 #keep changing alpha values
iterations <- 1000
c <- 0
m_vals <- c()
c_vals <- c()
error_vals <- c()

for (i in seq(1,iterations)){
  m_vals <- c(m_vals,m)
  c_vals <- c(c_vals,c)
  
  df <- mutate(df,mx_vals = m * x)              #formula for c_gradient
  df <- mutate(df, y_mx_vals_c = (y-mx_vals-c))
  c_gradient <- -2*sum(df$y_mx_vals_c)/nrow(df)
  
  df =  mutate(df, xy_vals = x * y)     #formula for m_gradient
  df = mutate(df, mx_square = m * (x^2))
  df = mutate(df,  cx_vals = c * x)
  df = mutate(df, total_m = xy_vals - mx_square - cx_vals)
  m_gradient = -2/nrow(df) * sum(df$total_m)
  
  df = mutate(df, current = (y - (m * x) - c)^2)       #to calculate error/current error
  error_vals <- c(error_vals, sum(df$current)/nrow(df))
  
  m = m - alpha * m_gradient
  c = c - alpha * c_gradient
  
}
print(c(m,c))


#install.packages("rgl")
library(rgl)
open3d()
plot3d(x=m_vals,y=c_vals,z=error_vals)



```









```{r}
mse = function(x,y,m,c){
  
  yhat=m*x+c
  error=sum((y-yhat)^2)/length(x)# error formula from book with yhat 
  return(error)

  }



cuts=100
c_ranges=seq(0,150,length.out = cuts)
m_ranges=seq(0,5,length.out = cuts)
zspace = c()
mspace = c()
cspace = c()

for(i in m_ranges){
  
  for(j in c_ranges){
    
  
    curr_z <- sum((df$y - i * df$x -j)^2) / nrow(df)
    zspace=c(zspace,curr_z)
    mspace = c(mspace, i)
    cspace = c(cspace, j)
    
  }
}
 
open3d()
plot3d(x=mspace,y=cspace,z=zspace,col = heat.colors(10))
plot3d(x=m_vals,y=c_vals,z=error_vals,add = T)


```



now taking a realtime dataset and applying gradient descent

here scaling x because the values of tv are very high


```{r}
adv <- read.csv("Advertising.csv")
df <- data.frame(x=scale(adv$TV),y=adv$sales)
alpha <- 1/nrow(df)
m <- 0
iterations <- 1000
c <- 0
m_vals <- c()
c_vals <- c()
error_vals <- c()

for (i in seq(1,iterations)){
  m_vals <- c(m_vals,m)
  c_vals <- c(c_vals,c)
  
  df <- mutate(df,mx_vals = m * x)              #formula for c_gradient
  df <- mutate(df, y_mx_vals_c = (y-mx_vals-c))
  c_gradient <- -2*sum(df$y_mx_vals_c)/nrow(df)
  
  df =  mutate(df, xy_vals = x * y)     #formula for m_gradient
  df = mutate(df, mx_square = m * (x^2))
  df = mutate(df,  cx_vals = c * x)
  df = mutate(df, total_m = xy_vals - mx_square - cx_vals)
  m_gradient = -2/nrow(df) * sum(df$total_m)
  
  df = mutate(df, current = (y - (m * x) - c)^2)       #to calculate error/current error
  error_vals <- c(error_vals, sum(df$current)/nrow(df))
  
  m = m - alpha * m_gradient
  c = c - alpha * c_gradient
  
}
print(c(m,c))


#install.packages("rgl")
library(rgl)
open3d()
plot3d(x=m_vals,y=c_vals,z=error_vals)

```






