---
title: "class10ML(recommendation system)"
author: "Suhail AK"
date: "May 22, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

topics to be covered

Recommendation System

Apriori algo
collaborative filtering
    - user based
    - item based
    
    
    
Packages needed for Apriori algo

```{r}
#install.packages("arules")
#install.packages("arulesViz")
library(reshape2)
#install.packages("recommenderlab")
library(arules)
library(arulesViz)
library(recommenderlab)

```

Apriori

```{r}
data(Groceries)
summary(Groceries)
class(Groceries)
dim(Groceries)
str(Groceries)

inspect(Groceries[1])


head(Groceries@itemInfo,n=30)
  
model <- apriori(data = Groceries,parameter = list(support=0.03,confidence=0.3))
inspect(model)


#Support: Support is the basic probability of an event to occur. If we have an event to buy product A, Support(A) is the number of transactions which includes A divided by total number of transactions.

#Confidence: The confidence of an event is the conditional probability of the occurrence; the chances of A happening given B has already happened.


```


```{r}

x <- list(c("a","b","c","d"),c("a","d"),c("e","a","c","b","d"),c("a","b","e","d"))
class(x)
x_trans <- as(x,'transactions')
x_trans
inspect(x_trans[1:4])
```


```{r}
model <- apriori(x_trans,parameter = list(support=0,confidence=0))
inspect(model)
```
support calculations
s(d) = f(D)/n
ie. f(D) is count and n is no. of transactions

```{r}
inspect(sort(model,by='lift',decreasing = T))
```

data of Groceries


```{r}
data(Groceries)
dim(Groceries)
model <- apriori(Groceries,parameter = list(support=(30/9835),confidence=0.5))
summary(model)

itemFrequencyPlot(Groceries,topN=20,main='Relative Item Frequency Plot',type="relative",ylab="Item Frequency (Relative)")

```

user based collaborative filtering

-cosine similarity
    find cosine similarity between user1 and user 2
    
    
working on ratings data

```{r}
library(reshape2)
library(tidyr)
rating <- read.csv("ratings.csv")
str(rating)
class(rating)
View(rating)
length(unique(rating$userId))# totally 671 users
length(unique(rating$movieId))#totally 9066 movies

ratings_matrix <- dcast(data=rating,userId~movieId, value.var = 'rating')

View(ratings_matrix[1:20,1:20])

dim(ratings_matrix)

rankings_matrix <- as(as.matrix(ratings_matrix[,-1]),'realRatingMatrix')

rankings_matrix@data

ubcf <- Recommender(rankings_matrix,method='UBCF', param=list(method='cosine',nn=10) )
#nn is no of neighbours ubcf user based collaboration

result <- predict(ubcf, rankings_matrix[5,],n=5)

movies_sugg <- as(result,'list')[[1]]

movies <- read.csv("movies.csv")

movies %>% filter(movieId %in% movies_sugg) %>% select(title)

View(rankings_matrix[1:10, 1:10])
```

Item based collaborative filtering.
IBCF: 

(movieid~userid)
method will have IBCF 
there is no nn in parameter list.


```{r}

rating_mat <- dcast(data =rating,movieId~userId,value.var = 'rating')
View(rating_mat[1:20,1:20])
dim(rating_mat)
rankings_matrix <- as(as.matrix(ratings_matrix[,-1]),'realRatingMatrix')

ibcf <- Recommender(rankings_matrix,method='IBCF', param=list(method='cosine') )

result <- predict(ubcf, rankings_matrix[1,],n=5)

movies_sugg <- as(result,'list')[[1]]

```

```{r}
data("Jester5k")
Jester5k
class(Jester5k)

ibcf <- Recommender(Jester5k,method='IBCF',parameter=list(method='cosine',k=10))

pred <- predict(ibcf,Jester5k[1], n=5)
as(pred,'list')

```

