---
title: "class9ml kmeans"
author: "Suhail AK"
date: "May 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


###k-means
- unsupervised learning
- input should be numerical
- if not then convert
- There is no training or testing concept


###Clustering(Unsupervised Learning)
- here there is no splitting of data into test and train
- feature selection is not necessary
- model used is kmeans

##working on iris data set

```{r}
library(MASS)
library(dplyr)
iri <- iris
str(iri)
nlevels(iri$Species)
iri <- iri %>% dplyr::select(-Species)
View(iri)

#to group using scatter plot

library(ggplot2)

ggplot(iri,aes(x = Petal.Length,y = Petal.Width))+geom_point()

#using kmeans

iris_model <- kmeans(iri,centers = 5)
iris_model
iris_model$cluster
table(iris_model$cluster) #length of each cluster
iris_model$size
iris_model$withinss

```

kmeans today using odi dataset

```{r}
odi <- read.csv("odi-batting.csv")
View(odi)
odi$century <- ifelse(odi$Runs>99,1,0)
odi$ducks <- ifelse(odi$Runs==0,1,0)
odi$above150 <- ifelse(odi$Runs>149,1,0)
odi$fifties <- ifelse(odi$Runs>49&odi$Runs<99,1,0)
odi$misses_century <- ifelse(odi$Runs>90& odi$Runs<99,1,0)
player_summary <- odi %>% group_by(Player) %>% summarise(matches=n(),
                                                         total_runs=sum(Runs,na.rm = T),
                                                         avg_runs=mean(Runs,na.rm = T),
                                                         centuries=sum(century,na.rm = T),
                                                         ducks=sum(ducks,na.rm = T),
                                                         fifties=sum(fifties,na.rm = T),
                                                         above_150=sum(above150,na.rm = T),
                                                        misses_century=sum(misses_century,na.rm = T))
View(player_summary)
```


building the model
while building remove all the categorical column
and normalize the data. using normalize function from library(class)

```{r}
top_players <- player_summary %>% arrange(-total_runs) %>% head(100)

data_kmeans <- top_players %>% dplyr::select(-Player)
data_norm <- normalize(data_kmeans,range = c(0,1),method = 'range')


model_kmeans <- kmeans(data_norm,centers = 5)
top_players$cluster <- model_kmeans$cluster
View(top_players)

barplot(table(top_players$cluster))
table(top_players$cluster)
model_kmeans$size

##total within
model_kmeans$tot.withinss

## withinsum of squares  this is distance of each clusters from the centroid , basically it should be less(close to each other)

model_kmeans$withinss


## between sum of squares (distance between 2 clusters hence this value should be ideally far or higher)

model_kmeans$betweenss



```



elbow method:  to find the optimal number of clusters.
plot with total_withinss as y axis
x-axis as no. of clusters





```{r}
View(data_norm)
dim(data_norm)
dim(as.matrix(dist(data_norm)))
data_norm_2d <- cmdscale(dist(data_norm)) #reducing the dimension
dim(data_norm_2d)
data_norm_2d <- as.data.frame(data_norm_2d)
plot(data_norm_2d)
data_norm_2d$cluster <- model_kmeans$cluster
data_norm_2d$cluster <- as.factor(data_norm_2d$cluster)

library(ggplot2)
ggplot(data_norm_2d,aes(x=V1,y=V2,color=cluster))+geom_point()

```


do the same for hr analytics
```{r}



```




##hierarchical clustering
good for small data set

```{r}
names(data_norm)
hclust_model <- hclust(dist(data_norm))
plot(hclust_model)
View(top_players)

```

### get cluster labels

```{r}
top_players <- player_summary %>% arrange(-total_runs) %>% head(100)
data_norm <- normalize(data_kmeans,range = c(0,1),method = 'range')
data_norm$cluster <- cutree(hclust_model,k = 4)
data_norm_2d <- cmdscale(dist(data_norm %>% dplyr::select(-cluster)))
data_norm_2d <- as.data.frame(data_norm_2d)
data_norm_2d$cluster <- as.factor(data_norm$cluster)
ggplot(data_norm_2d,aes(x=V1,y=V2,color=cluster))+geom_point()
View(data_norm_2d)

```


```{r}
install.packages("heatmaply")

cor_players <- cor(t(data_norm %>% head(10) %>% dplyr::select(-cluster)))
corrplot(cor_players,order = 'hclust',addrect = 2)


```

